Notes on scheduling the SAA2019 meeting

### 2017-11-04 analysis/paper/saa2018_posters.Rmd
posters: fill all day-parts will all posters, no overfill

@@

gen_poster_session_posters <- read_excel("../data/raw_data/02 - poster_sessions_FINAL(oct31).xlsx")

org_poster_sessions_only <- read_excel("../data/raw_data/02 - poster_sessions_FINAL(oct31).xlsx", sheet = "ORGANIZED_POSTER_SESSION_SIZES")

after_hours_poster_sessions_only <- 
  read_excel("../data/raw_data/02 - poster_sessions_FINAL(oct31).xlsx", 
             sheet = "AFTER_HOURS") %>% 
  slice(1:9) %>% 
  mutate(`Session Id` = as.numeric(gsub("\\D", "", ID)))

# In here we have org and after hours posters

@@

org_and_after_hours_posters_individual <- 
all_abstracts %>% 
  filter(`Session Type` == "Poster Symposium") %>% 
  filter(Role == "Presenter")

gen_and_org_poster_sessions_only <- 
all_abstracts %>% 
  filter(`Session Type` == "Poster Symposium") %>% 
  filter(Role == "Chair (Primary)")

# how many unique general sessions? 
n_gen_poster_sessions <- length(unique(gen_poster_session_posters$cuts))

# how many unique organised sessions? 
n_org_poster_sessions <- nrow((org_poster_sessions_only))

# how many unique after hours sessions? 
n_ah_poster_sessions <- nrow(after_hours_poster_sessions_only)

# total number of sessions
total_number_of_sessions <-  n_gen_poster_sessions + n_org_poster_sessions + n_ah_poster_sessions

# combine organised and general sessions
gen_poster_session_posters$session <- gen_poster_session_posters$cuts

org_poster_session_posters <- 
  org_and_after_hours_posters_individual %>% 
  mutate(session = `Session Id`) %>% 
  filter(session %in% org_poster_sessions_only$ID)

common_cols <- intersect(names(gen_poster_session_posters),  names(org_poster_session_posters))

gen_poster_session_posters <- 
  gen_poster_session_posters %>% 
  select(common_cols)

org_poster_session_posters <- 
  org_poster_session_posters %>% 
  select(common_cols)

gen_and_org_poster_session_posters <- 
  bind_rows(gen_poster_session_posters, 
            org_poster_session_posters)

# check the number of gen and org poster sessions:
gen_and_org_poster_sessions_only <- 
  gen_and_org_poster_session_posters %>% 
  group_by(session) %>% 
  slice(1)

after_hours_poster_sessions_all_posters <- 
  org_and_after_hours_posters_individual %>% 
  filter(`Session Id` %in% after_hours_poster_sessions_only$`Session Id`) %>% 
  mutate(poster_session_time = "Th-evening")

```

We have `r nrow(n_ah_poster_sessions)` sessions of after-hours posters, for a total of `r nrow(after_hours_poster_sessions_all_posters)` posters.

We have `r nrow(gen_and_org_poster_sessions_only)` sessions of general and organised session posters, for a total of `r nrow(gen_and_org_poster_session_posters)` posters.

@@

# add day-part to compare with papers
poster_session_durations <- 
  poster_session_durations %>% 
  mutate(start = as.character(start)) %>% 
  mutate(end = as.character(end)) %>% 
  mutate(day_part = case_when(
    start == "1899-12-31 08:00:00" ~ "morning",
    start == "1899-12-31 10:30:00" ~ "morning",
    start == "1899-12-31 14:00:00" ~ "afternoon",
    start == "1899-12-31 17:00:00" ~ "evening"
  )) %>% 
  mutate(day_part = glue('{stri_trans_totitle(substr(day, 1, 2))}-{day_part}'))

poster_session_durations
```

There are spaces for a total of `r sum(poster_session_durations$'N max')` posters. We have received `r nrow(gen_and_org_poster_session_posters)` poster submissions. So we have `r nrow(gen_and_org_poster_session_posters) - sum(poster_session_durations$'N max')` more posters than available spaces.

Thu-evening is the 'after hours' session, so that's already got the after-hours poster sessions in there. We'll just omit it from the following, and we don't need to worry about clashes because it's after all the papers. 100 posters are permitted in the after-hours session.

```{r}
# exclude Thurs evening from now on
poster_session_durations <- 
  poster_session_durations %>% 
  filter(day_part != "Th-evening")

# put out the after hours sessions and posters in a CSV
write_csv(after_hours_poster_sessions_all_posters,
          "../data/derived_data/after_hours_poster_sessions_all_posters.csv")

# Make sure that we don't have the Thu-evening posters in the remaining posterings
gen_and_org_poster_sessions_only$session[gen_and_org_poster_sessions_only$session %in% after_hours_poster_sessions_all_posters$`Session Id`] 
```

We have `r nrow(after_hours_poster_sessions_all_posters)` posters in the after hours session. That's `r ifelse(nrow(after_hours_poster_sessions_all_posters) <= 100, "great", "not ideal")`.

@@

```{r}
# prepare sessions we have
  poster_session_durations_summary <- 
    poster_session_durations %>% 
    group_by(day_part) %>% 
    summarise(`N max` = sum(`N max`))

  # make a unique value for names
  gen_and_org_poster_sessions_only <- 
  gen_and_org_poster_sessions_only %>% 
    mutate(uid = glue('{`First Name`}_{`Last Name`}'))

  gen_and_org_poster_session_posters <- 
    gen_and_org_poster_session_posters %>% 
    mutate(uid = glue('{`First Name`}_{`Last Name`}'))

  # we're going to change these
  gen_and_org_poster_sessions_only_tmp <- gen_and_org_poster_sessions_only
  container_for_this_day_part <- 0
  poster_session_durations_summary$numbers_of_posters <- 0
  poster_session_durations_summary$n_posters <- 0
  poster_session_durations_summary$n_sessions <- 0
  poster_session_durations_summary$c_sessions <- 0
  poster_session_durations_summary$c_posters <- 0
  all_of_the_day_parts <- 0
  i <- 1

# loop over each day-slot available for poster sessions
for(i in seq_len(nrow(poster_session_durations_summary))) {

    # stop loop gracefully when we have assigned all people to day-parts
  if (nrow(gen_and_org_poster_sessions_only_tmp) == 0){ break }

    # how many rows (posters) to we have remaning to allocate to a day-part?
  print(glue('number of poster sessions remaining to classify is {nrow(gen_and_org_poster_sessions_only_tmp)}'))

  # need a buffer
  buffer <- 3
  target_capacity <- poster_session_durations_summary$`N max`[i] - buffer

  # prepare variables for the next round, just to be sure
  cumulative_number_of_posters <- 0
  container_for_this_day_part <- 0
  this_session_posters <- 0
  j <- 1

  while(cumulative_number_of_posters <= target_capacity & j <= nrow(gen_and_org_poster_sessions_only_tmp)){

  # get a session 
  this_session <- gen_and_org_poster_sessions_only[j, ]

  # get names for all the people in this poster session
  this_session_posters <- 
    gen_and_org_poster_session_posters %>% 
      filter(session == this_session$session)

@@

      # check to see if papers for this day-part already have the name in it
      names_papers_in_this_day_part <- 
        sessions_in_day_parts_data_with_all_names_df %>% 
        filter(day_part_from_while.y == poster_session_durations_summary[i,]$day_part) %>% 
        pull(uid.y)

@@

#-------------------------------------------------------------------  
  # compute the number of posters we've added to this day-part
  # by counting the abstract Id, because this excludes the session chairs
    cumulative_number_of_posters <- 
      length(!is.na(container_for_this_day_part$`Abstract Id`))

   # store this result
    poster_session_durations_summary$numbers_of_posters[i] <- cumulative_number_of_posters

@@

   # add an identifier so we can see where we have assigned each session so far
 container_for_this_day_part_with_session_from_while <- 
   container_for_this_day_part %>% 
   mutate(day_part_from_while  = as.character(poster_session_durations_summary$day_part[i]))

@@

  poster_session_durations_summary$n_posters[i] <- length(unique(container_for_this_day_part_with_session_from_while$`Abstract Id`))

 poster_session_durations_summary$n_sessions[i] <- length(unique(container_for_this_day_part_with_session_from_while$session))

 poster_session_durations_summary$c_sessions[i] <-  sum(poster_session_durations_summary$n_sessions)
 poster_session_durations_summary$c_posters[i] <-  sum(poster_session_durations_summary$n_posters)

 print(poster_session_durations_summary %>% select(`N max`, 
                                                   day_part,
                                                   n_posters,
                                                   n_sessions, 
                                                   c_sessions,
                                                   c_posters))

@@

gen_and_org_poster_sessions_only_tmp <-  
   gen_and_org_poster_sessions_only_tmp %>% 
   filter(! session %in% unique(container_for_this_day_part$session))

@@

write.csv(poster_session_durations_summary, "../data/derived_data/poster_session_durations.csv")
```

Seems that we have too many posters for the available space. Let's double-check:

```{r}
# duplicates?
all_of_the_day_parts %>% 
  distinct()

>>>>

We can combine general and organised sessions, for the purpose of scheduling. BUt we have to get the after hours sessions and put them aside.

@@

# Here we get all the general session posters
# session number is cuts column
gen_poster_session_posters_individual <- read_excel("../data/raw_data/02 - poster_sessions_FINAL(oct31).xlsx")
gen_poster_session_posters_individual$`Session Id` <- gen_poster_session_posters_individual$cuts
# 422 posters in the general session

# grab the two general sessions from the 
# ah posters from the other sheet (very inconvenient!)
ah_poster_session_posters_individual <- read_excel("../data/raw_data/02 - poster_sessions_FINAL(oct31).xlsx", sheet = "AFTER_HOURS", skip = 11)
ah_poster_session_posters_individual$`Session Id` <- ah_poster_session_posters_individual$cuts

ah_poster_session_posters_individual <- 
ah_poster_session_posters_individual[, names(ah_poster_session_posters_individual) %in% names(gen_poster_session_posters_individual)  ]

gen_poster_session_posters_individual <- 
  rbind(gen_poster_session_posters_individual,
        ah_poster_session_posters_individual)

# In here we have org and after hours posters
# but not general posters

@@

org_and_after_hours_posters_individual <- 
all_abstracts %>% 
  filter(`Session Type` == "Poster Symposium") %>% 
  filter(Role == "Presenter")
# 196 posters in org + ah sessions

# combine general, organised and after hours individual posters
common_names <- 
names(gen_poster_session_posters_individual)[names(gen_poster_session_posters_individual) %in% names(org_and_after_hours_posters_individual)]

gen_poster_session_posters_individual <- 
  gen_poster_session_posters_individual %>% 
  select(common_names)

org_and_after_hours_posters_individual <- 
  org_and_after_hours_posters_individual %>% 
    select(common_names)

all_posters_individual <- 
  bind_rows(gen_poster_session_posters_individual,
            org_and_after_hours_posters_individual) %>% 
  filter(!is.na(Title))
```

We have `r nrow(all_posters_individual)` posters altogether, with `r nrow(gen_poster_session_posters_individual)` in the general session, and `r nrow(org_and_after_hours_posters_individual)` in the organised sessions (including after hours posters).

Let's remove the after hours sessions from all the posters, since we know they will be assigned to Th-evening:

```{r}
# how many ah sessions?
after_hours_poster_sessions_only <- 
  read_excel("../data/raw_data/02 - poster_sessions_FINAL(oct31).xlsx", 
             sheet = "AFTER_HOURS") %>% 
  slice(1:9) %>% 
  mutate(`Session Id` = as.numeric(gsub("\\D", "", ID)))

# how many ah posters?
after_hours_poster_sessions_posters_individual <- 
  all_posters_individual %>%
  filter(`Session Id` %in% after_hours_poster_sessions_only$`Session Id`) %>% 
  mutate(poster_session_time = "Th-evening")

# remove these from all posters
all_posters_individual_no_ah <- 
  all_posters_individual %>% 
  filter(!`Session Id` %in% after_hours_poster_sessions_only$`Session Id`)

# put out the after hours sessions and posters in a CSV
write_csv(after_hours_poster_sessions_posters_individual,
          "../data/derived_data/after_hours_poster_sessions_all_posters.csv")

# Make sure that we don't have the Thu-evening posters in the remaining posterings
all_posters_individual_no_ah$`Session Id`[all_posters_individual_no_ah$`Session Id` %in% after_hours_poster_sessions_posters_individual$`Session Id`] 
```

We have `r length(unique(after_hours_poster_sessions_only$'Session Id'))` sessions of after-hours posters, with a total of `r nrow(after_hours_poster_sessions_posters_individual)` posters. This leaves `r nrow(all_posters_individual_no_ah)` remaning to allocate in the day-parts.

100 posters are permitted in the after-hours session. We have `r nrow(after_hours_poster_sessions_posters_individual)` posters in the after hours session. That's `r ifelse(nrow(after_hours_poster_sessions_posters_individual) <= 100, "great", "not ideal")`.

@@

# add day-part to compare with papers
poster_session_durations <- 
  poster_session_durations %>% 
  mutate(start = as.character(start)) %>% 
  mutate(end = as.character(end)) %>% 
  mutate(day_part = case_when(
    start == "1899-12-31 08:00:00" ~ "morning",
    start == "1899-12-31 10:30:00" ~ "morning",
    start == "1899-12-31 14:00:00" ~ "afternoon",
    start == "1899-12-31 17:00:00" ~ "evening"
  )) %>% 
  mutate(start = substr(start, 11, nchar(start))) %>% 
  mutate(end = substr(end, 11, nchar(end))) %>% 
  mutate(day_part = glue('{stri_trans_totitle(substr(day, 1, 2))}-{day_part}'))

poster_session_durations

diff_between_need_and_have <- nrow(all_posters_individual_no_ah) - sum(poster_session_durations$'N max')
```

There are spaces for a total of `r sum(poster_session_durations$'N max')` posters. We have received `r nrow(all_posters_individual_no_ah)` poster submissions. So we have `r nrow(all_posters_individual_no_ah) - sum(poster_session_durations$'N max')` `r ifelse(diff_between_need_and_have < 0, glue('{abs(diff_between_need_and_have)} free spaces'), glue('{abs(diff_between_need_and_have)} more posters than available spaces'))`.

Thu-evening is the 'after hours' session, so that's already got the after-hours poster sessions in there. We'll just omit it from the following available schedules, and we don't need to worry about clashes because it's after all the papers.

And we'll combine day-parts that are in the same paper day-part, which will make it simper for avoiding clashes.

```{r}
# exclude Thurs evening from now on
poster_session_durations_no_th_ev <- 
  poster_session_durations %>% 
  filter(day_part != "Th-evening") %>% 
  group_by(day_part) %>% 
  summarise(n_max = sum(`N max`) )

poster_session_durations_no_th_ev
```

@@

```{r}
# to avoid overfilling day-parts
buffer <-7 # is the largest to avoid overfill

# day-parts to fill
 poster_session_durations_no_th_ev <- 
   poster_session_durations_no_th_ev %>% 
   arrange(desc(n_max))

# make a unique value for names for all posters
all_posters_individual_no_ah <- 
  all_posters_individual_no_ah %>% 
    mutate(uid = glue('{`First Name`}_{`Last Name`}'))

# get table of session id's only 
poster_session_ids <- 
unique(all_posters_individual_no_ah$`Session Id`)

# we're going to change these
  all_posters_individual_no_ah_tmp <- all_posters_individual_no_ah
  container_for_this_day_part <- 0
  poster_session_durations_no_th_ev$numbers_of_posters <- 0
  poster_session_durations_no_th_ev$n_posters <- 0
  poster_session_durations_no_th_ev$n_sessions <- 0
  poster_session_durations_no_th_ev$c_sessions <- 0
  poster_session_durations_no_th_ev$c_posters <- 0
  all_of_the_day_parts <- 0
  i <- 1

# loop over each day-slot available for poster sessions
# i is each day-part
for(i in seq_len(nrow(poster_session_durations_no_th_ev))) {

    # stop loop gracefully when we have assigned all people to day-parts
  if (nrow(all_posters_individual_no_ah_tmp) == 0){ break }

    # how many rows (posters) to we have remaning to allocate to a day-part?
  print(glue('number of poster sessions remaining to classify is {nrow(all_posters_individual_no_ah_tmp)}'))

  # need a buffer to avoid overfilling day-parts
  target_capacity <- poster_session_durations_no_th_ev$n_max[i] - buffer

  # prepare variables for the next round, just to be sure
  cumulative_number_of_posters <- 0
  container_for_this_day_part <- 0
  this_session_posters <- 0
  # j is each poster session
  j <- 1

  while(cumulative_number_of_posters <= target_capacity & 
        j <= length(poster_session_ids)) {

  # get a session 
  this_session <- poster_session_ids[j]

  # get names for all the people in this poster session
  this_session_posters <- 
    all_posters_individual_no_ah_tmp %>% 
      filter(`Session Id` == this_session)

@@

      # check to see if papers for this day-part already have the name in it
      names_papers_in_this_day_part <- 
        sessions_in_day_parts_data_with_all_names_df %>% 
        filter(day_part_from_while.y == poster_session_durations_no_th_ev[i,]$day_part) %>% 
        pull(uid.y)

@@

#-------------------------------------------------------------------  
  # compute the number of posters we've added to this day-part
  # by counting the abstract Id, because this excludes the session chairs
    cumulative_number_of_posters <- 
    nrow(container_for_this_day_part)

   # store this result
    poster_session_durations_no_th_ev$n_posters[i] <- cumulative_number_of_posters

@@

   # add an identifier so we can see where we have assigned each session so far
 container_for_this_day_part_with_session_from_while <- 
   container_for_this_day_part %>% 
   mutate(day_part_from_while  = as.character(poster_session_durations_no_th_ev$day_part[i]))

@@

 poster_session_durations_no_th_ev$n_sessions[i] <- length(unique(container_for_this_day_part_with_session_from_while$`Session Id`))

 poster_session_durations_no_th_ev$c_sessions[i] <-  sum(poster_session_durations_no_th_ev$n_sessions)
 poster_session_durations_no_th_ev$c_posters[i] <-  sum(poster_session_durations_no_th_ev$n_posters)

 print(poster_session_durations_no_th_ev %>% select(n_max, 
                                                   day_part,
                                                   n_posters,
                                                   n_sessions, 
                                                   c_sessions,
                                                   c_posters))

@@

all_posters_individual_no_ah_tmp <-  
   all_posters_individual_no_ah_tmp %>% 
   filter(! `Session Id` %in% unique(container_for_this_day_part$`Session Id`))

@@

write.csv(poster_session_durations_no_th_ev, "../data/derived_data/poster_session_durations_no_th_ev.csv")
```

We have from the SAA spaces for `r sum(poster_session_durations_no_th_ev$n_max)` posters. And we have `r sum(poster_session_durations_no_th_ev$c_posters)`

It's tricky to allocate the posters so they don't exceed the number allowed in each day part.

```{r}
# we have 532 posters not in ah
x1 <- 
all_posters_individual_no_ah

x2 <- 
all_of_the_day_parts

# what sessions are not in there?
session_output <- 
x2 %>% 
  group_by(`Session Id`) %>% 
  pull(`Session Id`) %>% 
  unique

session_input <- 
  x1 %>% 
  group_by(`Session Id`) %>% 
  pull(`Session Id`) %>% 
  unique

setdiff(session_input, session_output)

x1 %>% 
  filter(is.na(Title))

# what are in the output of this loop that is not in the input?
# duplicates?
all_of_the_day_parts %>% 
  distinct()
